### 2016-07-06 image_setup.tex
Edits on nodes configuration files section

\subsection{Configuring OS files and scripts}

In general, the \ac{Raspi}s should be setup as similar as possible. However,
some particularities exist to differentiate the devices in principle
One of those
few settings that may be different in the devices is their hostname.
%The hostname helps the user to distinguish the devices from each other.
%The hostname helps distinguish them from each other.
One approach to create unique hostnames would be to write the hostname
in each memory card. Another way is to create a start up script to automatically
configure the hostname. We decided to define the hostnames based on the \ac{MAC}
addresses of the \ac{Raspi}s wired ethernet interface. We store the \ac{MAC}
addresses and hostnames of the \ac{Raspi}s in a file as below. The testbed
administrator will need to insert the \ac{MAC} addresses and hostnames of
his/her \ac{Raspi}s.

%The \ac{Raspi}s should be setup to be as similar as possible. One of those
%things that is preferred to be different on the devices is their hostname.
%This helps distinguish them from each other. One solution to this in our
%approach would be to uniquely change the hostname in each memory card.
%This is however not desired because that would either require modifying
%the customized Raspbian lite image before writing the content to the
%memory card or to manually change it in the \ac{Raspi}s after the image has
%been written to a memory card.
%Instead, we decided to register the ethernet MAC addresses of all the \ac{Raspi}s
%in our testbed. Those had to be used to put them in a domain name anyway.
%
%Below is a part of the file where we keep track of the MAC addresses and assign
%hostnames. This file will be available online and copied to the Raspbian lite
%image in a later step under the listed filename. For different \ac{Raspi}s
%the MAC addresses has to be changed. The MAC address of a network card can be found
%using the command \texttt{"ifconfig"} or \texttt{"ip addr"}.

%Instead, we decided to register the ethernet MAC addresses of all the \ac{Raspi}s
%in our testbed. Those had to be used to put them in a domain name anyway.

%Below is a snippet of the file of all the hostnames and \ac{MAC} addresses:

%Below is a part of the file where we keep track of the MAC addresses and assign
%hostnames.
%This file will be available online and copied to the Raspbian lite
%image in a later step under the listed filename. For different \ac{Raspi}s
%the MAC addresses has to be changed. The MAC address of a network card can be found
%using the command \texttt{"ifconfig"} or \texttt{"ip addr"}.

% MAC and Hostname file
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/nodes.csv}@>
<@\textcolor{gray}{
---------------------------------------------------------------}
\Reactivatenumber @>
# Ethernet MAC    Hostname
b8:27:eb:5b:da:20 rasp00
b8:27:eb:7b:c3:91 rasp01
b8:27:eb:54:9c:64 rasp02
b8:27:eb:95:bd:11 rasp03
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

%Provided that each \ac{Raspi} will have knowledge of all other \ac{Raspi}s in
%the testbed, it still needs to know which name to give itself.
%We do this using a small Bash script:
%We use the \ac{Bash} script below to enable each \ac{Raspi} to find its
%hostname in \texttt{nodes.csv}.
A \ac{Bash} script can then be used to have each \ac{Raspi} assign its hostname:

% Set hostname
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/set\_hostname}@>
<@\textcolor{gray}{
---------------------------------------------------------------}
\Reactivatenumber @>
#!/usr/bin/env bash

@@

The code lines above do the following:
\begin{itemize}
    \item Line 1: Tells system to interpret the script using \ac{Bash}
    \item Line 3-4: Gets the path to the script itself and the list of hostnames
    \item Line 5: Gets the \ac{MAC} address of the node itself
    \item Line 6: Gets the current hostname
    \item Line 7: Gets the new hostname from the hostname list
    \item Line 10-14: Assigns the new hostname
\end{itemize}

%Line 1 tells the system to interpret the script using Bash. Line 3 gets the
%path of the script itself. Line 4 gets the MAC address. This script will be
%located in all \ac{Raspi}s and thus return the MAC address of the node itself
%when executed.
%Line 4 searches line by line for this mac address.
%Line 8-11 assigns the hostname found in \texttt{"nodes.csv"}. If the MAC did not exists,
%then it will keep the default hostname of the system.

>>>>

\subsection{Default OS Files and Configuration Scripts Setup}

In general, the \ac{Raspi}s should be setup as similar as possible. However,
some particularities exist to differentiate the devices in principle. Also,
scripts containing further configurations for the \ac{Raspi}s are desirable
to be distributed as part of the common image. Therefore, here we present
the steps to setup basic properties of the \ac{Raspi}s and distributing
configuration scripts to each of them through the image.

Among the settings that may be different in the devices is their hostname.
The hostname helps the user to physically distinguish the devices from
each other. Thus, we require the devices in our testbed to have different
hostnames. Hence, we define the hostnames based on the \ac{MAC} addresses
of the \ac{Raspi}s wired Ethernet interface. Prior to this stage, the
\ac{MAC} address of a network card can be found using the command
\texttt{ifconfig} or \texttt{ip addr} on a given \ac{Raspi}.
We store the \ac{MAC} addresses and hostnames of the \ac{Raspi}s in the
configuration file \texttt{\${ROOTDIR}/home/pi/rasp\_config/nodes.csv} as
shown:

% MAC and Hostname file
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/nodes.csv}@>
<@\textcolor{gray}{
--------------------------------------------------------------------------------}
\Reactivatenumber @>
# Ethernet MAC    Hostname
b8:27:eb:5b:da:20 rasp00
b8:27:eb:7b:c3:91 rasp01
b8:27:eb:54:9c:64 rasp02
b8:27:eb:95:bd:11 rasp03
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The testbed administrator has to insert the \ac{MAC} addresses and hostnames
of his/her \ac{Raspi}s obtained previously in the format shown in the
configuration file. Once we have defined the hostnames in
\texttt{\${ROOTDIR}/home/pi/rasp\_config/nodes.csv}, we use the following
\ac{Bash} script to have each \ac{Raspi} assign its hostname:

% Set hostname
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/home/pi/rasp\_config/set\_hostname}@>
<@\textcolor{gray}{
-------------------------------------------------------------------------------------}
\Reactivatenumber @>
#!/usr/bin/env bash

@@

The \texttt{\$\{ROOTDIR\}/home/pi/rasp\_config/set\_hostname} script (in
script lines): (1) tells the system to interpret the script using \ac{Bash},
(3-4) gets the path to the script itself and the list of hostnames, (5) gets
the \ac{MAC} address of the node itself, (6) gets the current hostname, (7)
gets the new hostname from the hostname list and (10-14) assigns the new
hostname to the \ac{Raspi} where the script will be executed.
